---
title: "Assignment 2: Coding Basics"
author: "Ruiqing Li"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<Li>_A02_CodingBasics.Rmd` (replacing `<Ruiqing>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics Day 1

1.  Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r}
#1. 
first_sequence <- seq(1, 100, 4) #Naming the sequence first_sequence

#2. 
mean(first_sequence) #Used sequence's assigned name first_sequence to get the mean
median(first_sequence) #Used sequence's assigned name first_sequence to get the median
#3. 
mean(first_sequence) > median(first_sequence) #Using Conditional Statement to compare mean of sequence and median sequence , the output (FALSE) means that the mean is not greater than the median. 
```

## Basics Day 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.

```{r}

Names <- c("Lisa","Cassie","Natalie","David") # Character Vectors
Names

TestScores <- c(23,48,92,99) #Numeric Vectors
TestScores

PassTestorNot <-c(FALSE,FALSE,TRUE,TRUE) #Logical Vectors

Grades <-data.frame(Names, TestScores, PassTestorNot)
Grades

```

9.  QUESTION: How is this data frame different from a matrix?

> Answer:Data frame can consist of multiple different classes of data instead of one.

10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. 

```{r}
#If and Else Function
Test_Score <- function(x){
  pass <- if (x>=50){
    TRUE
  }  else {
    FALSE
  }
  return(pass)
}
```

```{r}
#IfElse Function
Test_Score2 <- function(x){
  pass <- ifelse (x>=50,TRUE,FALSE)
  return(pass)
}
```

11. Apply your function to the vector with test scores that you created in number 5.


#If and Else Function
ifandelse<- Test_Score (TestScores)
ifandelse
**Error in if (x >= 50) { : the condition has length > 1**

```{r}
#IfElse Function
ifelse<- Test_Score2 (TestScores)
ifelse
```


12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer:`ifelse` function wroked. The vector TestScores have 4 different scores, `ifelse` function can run multiple numbers at a time where `if` and `else` function can only run a single number at a time. 
